Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-12-20 paper/main.tex
cleaning up latex compile errors + intro of pullback/pushforward

\section{Algebraic Topology \& Category Theory}
\note{potentially move this to appendix, if so push Butler & Vickers up to RW}

@@

The model described in this work provides a model for expressing data sets with different topological properties. 
We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
butlerVectorBundleClassesForm1992}. As described by them, a fiber bundle is a formal model of the mapping between data points and the underlying topological space they lie in. For example, nodes on a network and the graph of the network, an image and the underlying grid at which the image is sampled, or table columns and the table index.

@@

\subsection{Sheaves}
\label{sec:related-work:continuity:sheaves}
Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces. A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.

\begin{equation}
  \label{eq:related-work:continuity:presheaf}
  \begin{tikzcd}[column sep=small]
    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
    \end{tikzcd}
\end{equation}

This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$. Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dbase$ \cite{spanier1989algebraic, nlab:sheaf}. Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.

@@

 \begin{definition} A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a functor $\sheaf: \mathcal{C}^{op} \rightarrow \textbf{Set}$.\cite{nlab:presheaf } \end{definition} \note{possibly add in from Hader lecture notes}

@@

@@

For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase. For that line to have any thickness, it must be rendered as a ban. One method of obtaining this graphic continuity while preserving the constraint that it is a deformation retract is by multiplying the base space \dbase\ by an interval $I=\[0,1\]$ such that $\gbase = \dbase \times I^{1}$. This construction ensures that every region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to a  corresponding point $\dbasepoint \in \dbase$.

@@

When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance. This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}. The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$. The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$. The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

>>>>

\section{Algebraic Topology \& Category Theory}
\note{potentially move this to appendix, push Butler \& Vickers up to RW}

@@

The model described in this work provides a model for expressing data sets with different topological properties.

\subsection{Fiber Bundles}
We obtain this generality by using a mathematical structure called a fiber bundles as the basis of our abstraction, as proposed by Butler, Bryson, and Pendley\cite{butlerVisualizationModelBased1989,
butlerVectorBundleClassesForm1992}. As described by them, a fiber bundle is a formal model of the mapping between data points and the underlying topological space they lie in. For example, nodes on a network and the graph of the network, an image and the underlying grid at which the image is sampled, or table columns and the table index.

@@

@@

@@

\subsection{Sheaves}
\label{sec:related-work:continuity:sheaves}
Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets. Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces. A presheaf  $\sheaf$ applied to fiber bundles is a map from an open set in a topological space, such $\openset \subset \dbase$, to the set of sections 
\begin{equation}
  \label{eq:atct:presheaf}
  \sheafc: \opensetc \textcolor{sheaf}{\rightarrow} \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}
\end{equation}
over that openset \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves the inclusion maps $\iota$ between open sets and the restriction maps $\iota^*$ between the sets of sections over that open sets. 
\begin{equation}
  \label{eq:related-work:continuity:presheaf}
  \begin{tikzcd}[column sep=small]
    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook', color=set] \\& &\\
    \opensetc_1 \arrow[uu, "\sheafc_{\dtotalc}", color=sheaf] \arrow[rr, "\iota"', hook, color=base] & {} \arrow[uu, "\sheafc_{\dtotalc}" description, dotted, color=sheaf] & \opensetc_2 \arrow[uu, "\sheafc_{\dtotalc}"', color=sheaf]                    
    \end{tikzcd}
\end{equation}

This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$.

A presheaf \sheafc\, as introduced in \autoref{eq:related-work:continuity:presheaf}, is a contravariant functor\cite{nlab:presheaf} from a category to a set. 
\begin{equation}
  \label{eq:actc:sheaf}
 \sheafc: \textcolor{source}{\mathcal{C}^{op}} \rightarrow \textcolor{target}{\textbf{Set}}
\end{equation}

A presheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk\cite{StalkSheaf2019} 
\begin{equation}
  \label{eq:atct:stalk}
  \sheafc(\dbasec)\restriction_{\dbasepointc}\coloneqq \lim\limits_{\opensetc\ni \dbasepointc} \Gamma(\opensetc, \dtotalc\restriction_{\opensetc}) 
\end{equation}

Sheaves are presheafs where sections are continuous over all unions of open sets in $\dbase$\cite{harder2008lectures, spanier1989algebraic, nlab:sheaf}. This means that a section which is defined over a stalk $\dsection \in \sheaf(\dbase)\vert_{\dbasepoint_i}$ and evaluated at a point $\dbasepoint_j$ is equal to a section in the stalk over that point $\dsection_{dbasepoint_i}\vert_{dbasepoint_j} = \dsection_{\dbasepoint_j}$

We can define natural transforms, introduced in \autoref{eq:related-work:equivariance:natural-transform}, between sheaves over different spaces, for example spaces representing the topology of the data and the graphic, because sheaves can be moved over different topological spaces if there is a map between those spaces. 
Given a map $\textcolor{functor}{f}:\textcolor{source}{X} \textcolor{functor}{\rightarrow} \textcolor{target}{Y}$ between topological spaces $\textcolor{source}{X}, \textcolor{target}{Y}$ and sheaves on each space $\sheafc_{\textcolor{source}{X}}, \sheafc_{\textcolor{target}{Y}}$, we can construct the pullback functor $f^*$ and pushforward functor $f_*$ \cite{harder2008lectures} such that the following diagram commutes
\begin{equation}
  \label{eq:atct:pullback-pushforward}
  abc
\end{equation}

\note{possibly add in from Hader lecture notes}

@@

For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase. For that line to have any thickness, it must be rendered as a ban. One method of obtaining this graphic continuity while preserving the constraint that it is a deformation retract is by multiplying the base space \dbase\ by an interval $I=\[0,1\]$ such that $\gbase = \dbase \times I$. This construction ensures that every region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to a  corresponding point $\dbasepoint \in \dbase$.

@@

When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance. The fiber space over a point is embedded in the stalk, introduced in

over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$. The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$. The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.
